---
title: "JAMES issue resolution steps"
author: "Stef van Buuren"
format: html
---

```{r setup}
#| include: false
#| cache: false
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

## Objective

This document contains the step to create and deploy a new JAMES version.

## Local JAMES development

### Update and lock packages

Are all R packages up to data and complete?

```{r}
renv::status()
```

Use "explicit" mode to evade helpers like `devtools` and `roxygen2` in the lock file.

### Open, work and close issues in GitHub

-   Solve relevant issues in `james` package
-   <https://github.com/growthcharts/james/issues>

### Check the top level package `james`

```{r}
devtools::check()
```

### Update version files

If DESCRIPTION version changes, update two files with new version numbers

```{r}
james:::update_version_files("inst/www/index.html", 
                             "vignettes/articles/getting_started.Rmd", 
                             "DESCRIPTION")
```

### Update swaggerhub

- Free accounts no longer exist at SwaggerHub, but the last published version 1.5.4 is still current

### Next steps

-   Check `renv::status` and commit changes to `renv.lock`
-   Commit issue solution
-   Increase version
-   Update site: `devtools::build_site()`
-   Push changes to GitHub
-   Go to `jamesdocker` repo for building app
