---
title: "JAMES version maintaince"
author: "Stef van Buuren"
format: html
---

```{r setup}
#| include: false
#| cache: false
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

## Objective

- This document contains the steps to create a new JAMES version. 
- You can also `Render` this document to check and update the JAMES version in the package and documentation files.

## Local JAMES development

### Update and lock packages

Are all R packages up to data and complete?

```{r}
renv::status()
```

Use "explicit" mode to evade helpers like `devtools` and `roxygen2` in the lock file.

### Open, work and close issues in GitHub

-   Solve relevant issues in `james` package
-   <https://github.com/growthcharts/james/issues>

### Check the top level package `james`

```{r}
# devtools::check()
```

### Update the package version

-   Increase version number in DESCRIPTION file

### Update version in HTML and Rmd files

If DESCRIPTION version changes, update two files with new version numbers

```{r}
james:::update_version_files(
  "inst/www/index.html",
  "vignettes/articles/getting_started.Rmd",
  "DESCRIPTION"
)
```

### Update `openapi.yaml`

The source text of the OpenAPI specification is in `inst/spec/openapi.in.yaml`. Before building the package, this file needs updating with the current JAMES version from the `DESCRIPTION` file. This ensures that the two remain in sync.

```{r}
james:::update_openapi_spec()
```

### Keep `renv.lock` in shape

- Check `renv::status` and commit changes to `renv.lock`. 
- Only include packages that are relevant to the JAMES package, not the development packages in the project
- Use the function `james::install_dev_packages()` to install the development packages (like roxygen2) when starting fresh
- If needed, clean out the development packages by `renv::clean(prompt = TRUE)` from the library before using `renv::snapshot()` to update the lock file.

### Next steps

-   Update site: `devtools::build_site()`
-   Push changes to GitHub
-   Go to `jamesdocker` repo for building app
