---
title: "JAMES version maintaince"
author: "Stef van Buuren"
format: html
---

```{r setup}
#| include: false
#| cache: false
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

## Objective

- This document contains the steps to create a new JAMES version. 
- You can also `Render` this document to check and update the JAMES version in the package and documentation files.

## Local JAMES development

### Update and lock packages


- Use the function `james::install_dev_packages()` to install the development packages (like roxygen2) when starting fresh
- If needed, clean out the development packages by `renv::clean(prompt = TRUE)` from the library before using `renv::snapshot()` to update the lock file.


Are all R packages up to data and complete?

```{r}
renv::status()
```

If packages need updated, use `renv::restore()`.

### Open, work and close issues in GitHub

-   Solve relevant issues in `james` package
-   <https://github.com/growthcharts/james/issues>
-   Do relevant updates in underlying packages and install update `renv::install()`. Use github URL for github packages. Now project is out-of-sync, this is OK.
-   Update functions in `james` package that use these underlying functions
    -  (1) update james function
    -  (2) update documention `devtools::document()`

### Update `openapi.yaml`

The source text of the OpenAPI specification is in `inst/spec/openapi.in.yaml`. Before building the package, this file needs updating with the current JAMES version from the `DESCRIPTION` file. This ensures that the two remain in sync.

To add an argument to dcat, search for dcat_arguments and add the argument that was added to the function. 


### Check the top level package `james`

```{r}
# devtools::check()
```

### Update the package version

-   Update `NEWS.md` with version update and document changes.
-   Increase version number in DESCRIPTION file
-   Also update imports version with proper versions for updated packages.

```{r}
# devtools::check()
```


### Update version in HTML and Rmd files

If DESCRIPTION version changes, update two files with new version numbers

-   Rename `getting_started.qmdx` --> `getting_started.qmd`

```{r eval= FALSE}
james:::update_version_files(
  "inst/www/index.html",
  "vignettes/articles/getting_started.qmd",
  "DESCRIPTION"
)
```

Update version number van openapi.in.yaml file

```{r eval=FALSE}
james:::update_openapi_spec()
```


### Keep `renv.lock` in shape

- Check `renv::status` and commit changes to `renv.lock`. 
- Only include packages that are relevant to the JAMES package, not the development packages in the project

Use "explicit" mode to evade helpers like `devtools` and `roxygen2` in the lock file. 
```{r}
renv::snapshot(type = "explicit")
```


### Next steps


-   RUN in terminal: Rscript -e "quarto::quarto_render(input = 'vignettes/articles/getting_started.qmd')" - does not work in Windows because issues with bash codes
-   By default, this script runs server "test" (james.groeidiagrammen.nl), which initially contains the previous JAMES version
-   We first have to build and test our new version on localhost and test

-   Rename `getting_started.qmd` --> `getting_started.qmdx` to exempt getting_started from being build by `pkgdown::build_site()`

-   Update site: `pkgdown::build_site()`
-   Move `vignettes/articles/getting_started.html` to `docs/articles/getting_started.html`
-   Move `vignettes/articles/geting_started_files/` to `docs/articles/getting_started_files/`

-   Push changes to GitHub in following separate commits
    - (1) function changes
    - (2) update package (news description)
    - (3) documentation
-   Go to `jamesdocker` repo for building app
