---
title: "JAMES version maintaince"
author: "Stef van Buuren"
format: html
---

```{r setup}
#| include: false
#| cache: false
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

## Objective

- This document contains the recipe to create a new JAMES version.
- Follow directions sequentially
- Commit where indicated
- Manually run the code blocks
- Inspect the results after each step before progressing

## Local JAMES development

## STEP 1 - FUNCTIONALITY

### Update and lock packages

- Use the function `james::install_dev_packages()` to install the development packages (like roxygen2) when starting fresh
- If needed, clean out the development packages by `renv::clean(prompt = TRUE)` from the library before using `renv::snapshot()` to update the lock file.

Are all R packages up to date and complete?

```{r eval=FALSE}
renv::status()
```

If packages need updating, use `renv::restore()`. Then snapshot using "explicit" mode to evade helpers like `devtools` and `roxygen2` in the lock file.

```{r eval=FALSE}
renv::snapshot(type = "explicit")
```

### Open, work and close issues in GitHub

-   Solve relevant issues in `james` package
-   <https://github.com/growthcharts/james/issues>
-   Do relevant updates in underlying packages and install with `renv::install()`. Use GitHub URL for GitHub packages. Now project is out-of-sync, this is OK.
-   Update functions in `james` package that use these underlying functions
    -  (1) update james function
    -  (2) update documentation `devtools::document()`

### Update james ecosystem version numbers and functionality for deployment

- Update imports version with proper versions for updated packages.
- Update new/deprecated functions, arguments, examples etc in the OpenAPI specification at `inst/spec/openapi.in.yaml`. For example, to add a new argument to `dcat`, search for `dcat_arguments` and add the argument to the description. Be careful to use proper spacing and syntax. YAML is unforgiving.

### Check the top level package `james`

Run `devtools::check()` and fix any errors or warnings before proceeding.

```{r eval=FALSE}
devtools::check()
```

### COMMIT 1

Commit 1: function changes, `devtools::document()` output, `renv.lock`, `inst/spec/openapi.in.yaml`, `vignettes/articles/_getting_started.qmd`

## STEP 2 - VERSIONING

### Update JAMES package version and NEWS

- Update `NEWS.md` with version update and document changes.
- Increment the `DESCRIPTION` version number, e.g.
  + for a minor release 1.13.x --> 1.13.{x+1}
  + for a new docker version 1.13.x --> 1.14.0

### Sync `openapi.yaml`

Sync the version number in `inst/www/openapi.in.yaml` from `DESCRIPTION` using

```{r eval=FALSE}
james:::update_openapi_spec()
```

### Sync `index.html`, `_getting_started.qmd`

Sync version numbers from `DESCRIPTION` using

```{r eval=FALSE}
james:::update_version_files(
  "inst/www/index.html",
  "vignettes/articles/_getting_started.qmd",
  "DESCRIPTION"
)
```

### COMMIT 2

`NEWS.md`, `DESCRIPTION`, `inst/www/openapi.yaml`, `vignettes/articles/_getting_started.qmd`, `inst/www/index.html`


## STEP 3 - DOCUMENTATION

### Update documentation

- Build documentation using a combination of `pkgdown::build_site()` and `quarto::quarto_render()`
- `quarto_render()` builds the getting started vignette
- `bash` code does not work on Windows, and will be skipped by setting `eval=FALSE` under Windows
- By default, the script runs server `target_host = "test"` (james.groeidiagrammen.nl)
- Note: `getting_started.qmd` also runs within the container, so check any changes in the `jamesdocker` package

Run the following chunk to update the version in source files, render the vignette, and build the site:

```{r eval=FALSE}
# Step 1: render _getting_started.qmd to docs/articles/
quarto::quarto_render(
  input = 'vignettes/articles/_getting_started.qmd',
  quarto_args = c('--output-dir', file.path(getwd(), 'docs/articles'))
)

# Step 2: rename _getting_started.html --> getting_started.html
file.rename(
  'docs/articles/_getting_started.html',
  'docs/articles/getting_started.html'
)

# Step 3: rename _getting_started_files/ --> getting_started_files/
from <- "docs/articles/_getting_started_files"
to <- "docs/articles/getting_started_files"
ok <- file.rename(from, to)
if (!ok) {
  dir.create(to, recursive = TRUE, showWarnings = FALSE)
  file.copy(
    from = list.files(from, full.names = TRUE, all.files = TRUE, no.. = TRUE),
    to = to,
    recursive = TRUE,
    overwrite = TRUE
  )
  unlink(from, recursive = TRUE)
}

# Step 4: fix internal references in getting_started.html
html <- readLines('docs/articles/getting_started.html')
html <- gsub(
  '_getting_started_files/',
  'getting_started_files/',
  html,
  fixed = TRUE
)
writeLines(html, 'docs/articles/getting_started.html')

# Step 5: render dcat_calculate_examples.qmd to docs/articles/
quarto::quarto_render(
  input = 'vignettes/articles/dcat_calculate_examples.qmd',
  quarto_args = c('--output-dir', file.path(getwd(), 'docs/articles'))
)

# Step 6: build pkgdown site
pkgdown::build_site()
```

### COMMIT 3

`*.qmd`, `docs/`


## Push and go

-   Push the commits to GitHub
-   Go to `jamesdocker` repo for building app
