---
output:
  md_document:
    variant: markdown_github
bibliography: [README.bib]
---

<!-- README.md is generated from README.Rmd. Please edit that file -->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval=FALSE)
```

## james

The `james` package implements the **Joint Anthropometric Measurement
and Evaluation System (JAMES)**, an **experimental** online resource
for creating growth charts.

The webservice can be used by anyone interested in high-quality charts for
monitoring and evaluating childhood growth.

This document explains how to set up communications with JAMES and
provides some guidance in its use.

## Architecture

The growth charts in JAMES are programmed in `R`. JAMES makes these
available through the [OpenCPU](https://www.opencpu.org) system for
scientific computing and reproducuble research. The system allows for
easy integration of growth charts into any `HTTP` compliant client by
means of OpenCPU's [API](https://www.opencpu.org/api.html). JAMES is 
a RESTful webservice.

## Access

JAMES is currently located at url `groeidiagrammen.nl`. The
sections below use `curl` for illustration, but any `HTTP` client will
work.

In order to check whether the JAMES server is running, try generating some random 
numbers by calling `stats::rnorm()`, in a terminal window as follows,

```{bash}
curl http://groeidiagrammen.nl/ocpu/library/stats/R/rnorm/json --data n=5

[2.259, 0.0538, 1.0583, 0.8665, 0.8412]
```

## What JAMES can currently do

JAMES can currently do four things:

1. provides access to 342 high-quality growth charts used by the Dutch youth health care;
2. can accept child data in BDS-format;
3. can create a personalized chart site with child's data;
4. can create separate growth charts.

I'll explain each of these action in more detail below.

## Available growth charts 

Copy the [following url](http://groeidiagrammen.nl/ocpu/lib/james/www/) into your browser to obtain an interactive overview of the available growth charts:

```{bash}
http://groeidiagrammen.nl/ocpu/lib/james/www/
```

There are 342 different charts, each identified with a `chartcode`, 
a 4-7 character code identifying each design. The `list_charts()` function in 
JAMES function can produce a tabular overview. Obtain its help file as

```{bash}
curl http://groeidiagrammen.nl/ocpu/lib/james/man/list_charts/text
```

The list of growth charts can be created by the following request:

```{bash}
curl http://groeidiagrammen.nl/ocpu/lib/james/R/list_charts -d ""

/ocpu/tmp/x045d183eddf2f4/R/.val
/ocpu/tmp/x045d183eddf2f4/R/list_charts
/ocpu/tmp/x045d183eddf2f4/stdout
/ocpu/tmp/x045d183eddf2f4/source
/ocpu/tmp/x045d183eddf2f4/console
/ocpu/tmp/x045d183eddf2f4/info
/ocpu/tmp/x045d183eddf2f4/files/DESCRIPTION
```

which return a list of 8 urls created by `OpenCPU`. The table can be 
download by

```{bash}
curl http://groeidiagrammen.nl/ocpu/tmp/x045d183eddf2f4/R/.val/print

    chartgrp chartcode population    sex design  side language week
1     nl2010      HJAA         HS   male      A front    dutch
2     nl2010      HJAB         HS   male      A  back    dutch
3     nl2010      HJAH         HS   male      A   hgt    dutch
4     nl2010      HJAO         HS   male      A   hdc    dutch
...
```

The chart codes are compatible with the codes used in @talma2010. 

Note that the session key `x045d183eddf2f4` in the above example is
specific to the call, and will be different for each session. Session
keys and their url's on the JAMES server will be removed after 24 
hours.

## Posting BDS data to JAMES

### Sending the child's data as a JSON file

Suppose that the child's data are coded according to the specification
[BDS JGZ 3.2.5](https://www.ncj.nl/themadossiers/informatisering/basisdataset/documentatie/?cat=12),
and converted into `JSON` format. The JAMES server contains the
following example:

```{bash}
curl http://groeidiagrammen.nl/ocpu/library/james/testdata/client3.json
```

Save the file
[client3.json](http://groeidiagrammen.nl/ocpu/library/james/testdata/client3.json)
on your local system by

```{bash eval=FALSE}
curl http://groeidiagrammen.nl/ocpu/library/james/testdata/client3.json -O
```

which produces a file called `client3.json` in your work directory. 

For testing purposes, you may change its values (but please keep the
general structure intact) and re-upload it to JAMES for analysis. 

The following `curl` command re-uploads and converts the file
`client3.json` from BDS format into an `R` object of S4 class
`minihealth::individual`:

```{bash, eval = FALSE}
curl -F 'txt=@client3.json' http://groeidiagrammen.nl/ocpu/library/james/R/convert_bds_ind

/ocpu/tmp/x06938035d05dac/R/.val
/ocpu/tmp/x06938035d05dac/R/convert_bds_ind
/ocpu/tmp/x06938035d05dac/stdout
/ocpu/tmp/x06938035d05dac/source
/ocpu/tmp/x06938035d05dac/console
/ocpu/tmp/x06938035d05dac/info
/ocpu/tmp/x06938035d05dac/files/client3.json
/ocpu/tmp/x06938035d05dac/files/DESCRIPTION
```

An print version of the result is stored in 

```{r}
curl http://groeidiagrammen.nl/ocpu/tmp/x06938035d05dac/R/.val/print

An object of class "individual"
Slot "child":
data frame with 0 columns and 0 rows

Slot "time":
data frame with 0 columns and 0 rows

Slot "id":
[1] 0

Slot "name":
[1] "fa308134-069e-49ce-9847-ccdae380ed6f"

Slot "dob":
[1] "11-10-18"

Slot "src":
[1] "1234"

Slot "dnr":
[1] NA

Slot "sex":
[1] "female"

Slot "etn":
[1] "NL"

Slot "edu":
[1] NA

Slot "ga":
[1] 27

Slot "bw":
[1] 990

Slot "twin":
[1] NA

Slot "agem":
[1] NA

Slot "smo":
[1] 0

Slot "hgtm":
[1] 1670

Slot "wgtm":
[1] NA

Slot "hgtf":
[1] 1900

Slot "wgtf":
[1] NA

Slot "bfexc06":
[1] NA

Slot "durbrst":
[1] NA

Slot "hgt":
package: clopus, library: clopus::nl1997 , member: nl1997.fhgtNL 
     age  hgt     hgt.z
1 0.0849 38.0 -7.943717
2 0.1670 43.5 -6.385568

Slot "wgt":
package: clopus, library: clopus::nl1997 , member: nl1997.fwgtNL 
     age  wgt     wgt.z
1 0.0849 1.25 -8.583512
2 0.1670 2.10 -6.581767

Slot "hdc":
package: clopus, library: clopus::nl1997 , member: nl1997.fhdcNL 
     age  hdc     hdc.z
1 0.0849 27.0 -7.548492
2 0.1670 30.5 -6.066642

Slot "bmi":
package: clopus, library: clopus::nl1997 , member: nl1997.fbmiNL 
     age      bmi     bmi.z
1 0.0849  8.65651 -5.719143
2 0.1670 11.09790 -3.766751

Slot "wfh":
No reference
   hgt  wfh wfh.z
1 38.0 1.25    NA
2 43.5 2.10    NA

Slot "bs.hgt":
package: donordata, model: donordata::smocc_bs , member: hgt 
      age      hgt     hgt.z
1  0.0000 37.63782 -6.929621
2  0.0767 38.73560 -7.445715
3  0.1533 42.68033 -6.591987
4  0.2500 46.91763 -5.895737
5  0.3333 50.40139 -5.332599
6  0.5000 55.73412 -4.619819
7  0.6250 58.76424 -4.255843
8  0.7500 61.29885 -4.003211
9  0.9167 63.95774 -3.879229
10 1.1667 67.36804 -3.820196
11 1.5000 71.67160 -3.603667
12 2.0000 77.18910 -3.270500
13 3.0000 90.09290 -1.771377

Slot "bs.wgt":
package: donordata, model: donordata::smocc_bs , member: wgt 
      age        wgt     wgt.z
1  0.0000  0.9056132 -9.006665
2  0.0767  1.2496280 -8.454686
3  0.1533  1.9393741 -6.894393
4  0.2500  2.8894599 -5.392084
5  0.3333  3.5923136 -4.653004
6  0.5000  4.9277774 -3.525314
7  0.6250  5.5292387 -3.306713
8  0.7500  6.1150749 -3.054297
9  0.9167  6.6887088 -2.946767
10 1.1667  7.3461849 -2.920375
11 1.5000  8.2449663 -2.680337
12 2.0000  9.2358047 -2.608811
13 3.0000 10.7343002 -2.773075

Slot "bs.hdc":
package: donordata, model: donordata::smocc_bs , member: hdc 
      age      hdc     hdc.z
1  0.0000 24.74355 -8.185195
2  0.0767 27.30185 -7.188857
3  0.1533 30.11832 -6.148961
4  0.2500 32.12195 -5.860350
5  0.3333 34.30669 -5.121962
6  0.5000 36.13303 -5.082818
7  0.6250 37.72714 -4.571027
8  0.7500 38.47569 -4.519543
9  0.9167 39.62832 -4.234684
10 1.1667 40.34018 -4.247731
11 1.5000 41.79837 -3.710716
12 2.0000 42.51795 -3.634945
13 3.0000 43.21616 -3.808453

Slot "bs.bmi":
package: donordata, model: donordata::smocc_bs , member: bmi 
      age       bmi     bmi.z
1  0.0000  9.191473 -3.230867
2  0.0767  9.055411 -5.031675
3  0.1533 10.752560 -4.023688
4  0.2500 12.188298 -3.201624
5  0.3333 13.524967 -2.264696
6  0.5000 13.881065 -2.343775
7  0.6250 13.829450 -2.527176
8  0.7500 14.487793 -1.969734
9  0.9167 14.597060 -1.853467
10 1.1667 14.419768 -1.878160
11 1.5000 14.525894 -1.568160
12 2.0000 14.376370 -1.485082
13 3.0000 13.781281 -1.783393

Slot "bs.wfh":
Broken stick model not found.
  hgt wfh wfh.z
1  NA  NA    NA
```

The child's data on the server may be used for plotting on growth charts.

### Sending the child's data as a JSON string

If desired, the child's data may be sent by a JSON string instead of 
a file. The following code assumes that utility `jq` is installed.

First, make the json file into a JSON string, with double quotes (`"`)
properly escaped.

```{bash}
var=$(jq '.' client3.json | jq -sR '.')
echo $var
```

Then paste into into the call to the OpenCPU server: 

```{bash}
curl http://groeidiagrammen.nl/ocpu/library/james/R/convert_bds_ind -d "txt=$var"
```

The results of the string- and file-methods are identical.

## Request a dedicated url to the chart site

The child's data can be plotted on any of the available charts using 
a *child chart site*. There are three ways to choose which chart is 
being plotted:

1. JAMES picks a default chart;
2. The developer provides a `chartcode` parameter;
3. The end user manipulates the interactive controls.

Options 1 and 2 determine the first chart that shown to the end user.

### Site with default chart

The default chart picked by JAMES is currently hard-wired as the
child's height chart that contains the most recent measurements. If
the child is a pre-term (gestational age <= 36 weeks) and younger than
4 years, then JAMES chooses the appropriate preterm chart.

The chart site with the default start can be started by combining the 
[uploaded data](http://groeidiagrammen.nl/ocpu/tmp/x06938035d05dac/)
and the main site at (http://groeidiagrammen.nl/ocpu/lib/james/www/) 
as
```{bash}
curl "http://groeidiagrammen.nl/ocpu/lib/james/www/?ind=http://groeidiagrammen.nl/ocpu/tmp/x06938035d05dac/"
```

Pasting this url in your browser starts the site with the child's data.

### Site with developer-specified chart

Starting the site at a given growth chart is possible by specifying the 
`chartcode` parameters. For example, we may initialize the site at chart 
`PMAAN27` by 

```{bash}
curl "http://groeidiagrammen.nl/ocpu/lib/james/www/?ind=http://groeidiagrammen.nl/ocpu/tmp/x06938035d05dac/&chartcode=PMAAN27"
```

The site now starts with `PMAAN27` instead of `PMAHN27`. Any chart can
be chosen. It is the responsability of the developer that the choice
is sensible given the child's data.
 
### Site with user-specified chart

After the site is started the end user may change the chart on which the
data are drawn by simply using the site controls.

## Requesting a single growth chart

Requesting a single growth chart can be done through the main JAMES
chart drawing function is `draw_chart()`. 

The `draw_chart()` function does not yet support the creation of single 
graphs. The `draw_chart_bds()` and `draw_chart_ind()` are temporary 
specialty functions that can do this. The functions are documented
in

```{bash eval=TRUE}
curl http://groeidiagrammen.nl/ocpu/lib/james/man/draw_chart/text
```

Use `draw_chart_bds()` to draw chart directly from the BDS data file.

```{r}
curl -F 'txt=@client3.json' http://groeidiagrammen.nl/ocpu/library/james/R/draw_chart_bds

/ocpu/tmp/x0a862a7b6829a1/R/.val
/ocpu/tmp/x0a862a7b6829a1/R/draw_chart_bds
/ocpu/tmp/x0a862a7b6829a1/graphics/1
/ocpu/tmp/x0a862a7b6829a1/source
/ocpu/tmp/x0a862a7b6829a1/console
/ocpu/tmp/x0a862a7b6829a1/info
/ocpu/tmp/x0a862a7b6829a1/files/client3.json
/ocpu/tmp/x0a862a7b6829a1/files/DESCRIPTION
```

Download the growth chart as

```{bash}
curl -o mychart.svg http://groeidiagrammen.nl/ocpu/tmp/x0a862a7b6829a1/graphics/1/svg
```

or view it in the browser by paste the following into the url address field:

```{bash}
http://groeidiagrammen.nl/ocpu/tmp/x0a862a7b6829a1/graphics/1/svg
```

If the data are already uploaded, we may use `draw_chart_ind()`. Suppose 
we want to create the A4 chart for the child. We can do so by

```{bash}
curl http://groeidiagrammen.nl/ocpu/lib/james/R/draw_chart_ind -d "ind_loc='http://groeidiagrammen.nl/ocpu/tmp/x06938035d05dac/'&chartcode='PMAAN27'"

/ocpu/tmp/x029e4ee555a9e1/R/.val
/ocpu/tmp/x029e4ee555a9e1/R/draw_chart_ind
/ocpu/tmp/x029e4ee555a9e1/graphics/1
/ocpu/tmp/x029e4ee555a9e1/source
/ocpu/tmp/x029e4ee555a9e1/console
/ocpu/tmp/x029e4ee555a9e1/info
/ocpu/tmp/x029e4ee555a9e1/files/DESCRIPTION
```

where we used the data previously stored from session `x06938035d05dac`.

As before, we may browse the graph by pasting the following in the address field of the browser: 

```{bash}
http://groeidiagrammen.nl/ocpu/tmp/x029e4ee555a9e1/graphics/1/svg?width=8.27&height=11.69
```

or download it the the client's system.

The functions `draw_chart_bds()` and `draw_chart_ind()` address
special cases. When they are integrated in future versions of
`draw_chart()`, both functions will be deprecated.

## A client in `R`

If `R` is your analysis environment, then you can use the
`james.client` package to achieve the same result in an easier way.
The functions `upload_bds()` and `request_chart()` in `james.client`
have simple syntax, and automate the steps outlined above. See
<https://github.com/stefvanbuuren/james.client> for more detail.

## Next steps

Still many things on the wish list:

- transfer JAMES to url james.tno.nl
- allow for more input formats
- add https protocol
- add functionality to test for Dutch guidelines for referral
- add functionality to predict individual growth curves
- extend functionality to include the $D$-score charts

## Known problems

- Older versions of Edge (before 44.11763.1.0) do not display charts

## Resources

- [OpenCPU system](https://www.opencpu.org)
- [OpenCPU API](https://www.opencpu.org/api.html)
- https://www.w3schools.com/js/
- https://www.tno.nl/groei and https://www.tno.nl/growth
- https://github.com/stefvanbuuren/james
- https://github.com/stefvanbuuren/james.client
- https://github.com/stefvanbuuren/minihealth
- https://github.com/stefvanbuuren/brokenstick

## About

**Work in progress**. Direct suggestions and inquiries to Stef van Buuren
(stef.vanbuuren at tno.nl), <https://stefvanbuuren.name>, <https://github.com/stefvanbuuren>.

## Literature

