---
title: "JAMES dcat/calculate endpoint"
author: "Stef van Buuren"
date: today
format:
  html:
    number-sections: false
    toc: true
    toc-depth: 3
    code-copy: true
    code-tools: true
    theme: united
    embed-resources: true
---

```{r setup}
#| include: false
library(httr2)
library(jsonlite)
is_windows <- .Platform$OS.type == "windows"
host <- "https://james.groeidiagrammen.nl"
knitr::opts_chunk$set(python.reticulate = FALSE) # skip python execution
```

## Overview

The `/dcat/calculate` endpoint performs adaptive testing of developmental
milestones. Given child data in BDS format 3.1 it returns the next milestone
item to administer, together with supporting D-score information.

The minimal input is the child's date of birth (`bdsNumber: 20`). Prior
milestone responses can be included as `clientMeasurements`.

**Host used in these examples**: `https://james.groeidiagrammen.nl` (open
access, no authentication required).

**Notes**

- `?auto_unbox=true` unwraps single-element JSON arrays in the response —
  without it, scalars come back as `[value]`.
- ACC and PROD servers require a Bearer token added to the request headers.
- The default instrument is `gs1`; pass `instrument = "ddi"` (plus a
  compatible `key`) to use DDI items instead.
- These examples use `population = "GSED-NLD"`, `p = 90` (percentile threshold)
  and `k_domain = 3` (number of domains).

## Examples

::: panel-tabset

### R

Install and load the required packages once:

```r
install.packages(c("httr2", "jsonlite"))
remotes::install_github("growthcharts/bdsreader")
```

POST child data from a file (default `gs1` instrument):

```{r dcat-r-file}
fn <- system.file("examples", "example_v3.1.json", package = "bdsreader")

resp <- request(paste0(host, "/dcat/calculate/json")) |>
  req_url_query(auto_unbox = "true") |>
  req_body_multipart(
    txt = curl::form_file(fn, type = "application/json"),
    population = "\"GSED-NLD\"",
    p = "\"90\"",
    k_domain = "\"3\""
  ) |>
  req_headers(accept = "application/json") |>
  req_perform()

resp_body_json(resp)
```

POST a JSON object built inline — written to a tempfile for upload:

```{r dcat-r-inline}
child_data <- list(
  Format = "3.1",
  clientDetails = list(list(bdsNumber = 20L, value = "20260115"))
)

tmp <- tempfile(fileext = ".json")
write_json(child_data, tmp, auto_unbox = TRUE)

resp <- request(paste0(host, "/dcat/calculate/json")) |>
  req_url_query(auto_unbox = "true") |>
  req_body_multipart(
    txt = curl::form_file(tmp, type = "application/json"),
    population = "\"GSED-NLD\"",
    p = "\"90\"",
    k_domain = "\"3\""
  ) |>
  req_headers(accept = "application/json") |>
  req_perform()

resp_body_json(resp)
```

Include a prior `gs1` milestone response — JAMES uses it to select the next adaptive item:

```{r dcat-r-rich}
child_data <- list(
  Format = "3.1",
  clientDetails = list(list(bdsNumber = 20L, value = "20260115")),
  clientMeasurements = list(
    list(
      varName = "gs1moc002",
      values = list(list(date = "20260504", value = "1"))
    )
  ),
  nestedDetails = list()
)

tmp <- tempfile(fileext = ".json")
write_json(child_data, tmp, auto_unbox = TRUE)

resp <- request(paste0(host, "/dcat/calculate/json")) |>
  req_url_query(auto_unbox = "true") |>
  req_body_multipart(
    txt = curl::form_file(tmp, type = "application/json"),
    population = "\"GSED-NLD\"",
    p = "\"90\"",
    k_domain = "\"3\""
  ) |>
  req_headers(accept = "application/json") |>
  req_perform()

resp_body_json(resp)
```

### bash

POST a local file:

```{bash dcat-bash-file, eval=!is_windows}
HOST="https://james.groeidiagrammen.nl"
FN=$(Rscript -e 'cat(system.file("examples", "example_v3.1.json", package = "bdsreader"))')

curl -sX POST "$HOST/dcat/calculate/json?auto_unbox=true" \
  -H "accept: application/json" \
  -F "txt=@$FN;type=application/json" \
  -F "population='GSED-NLD'" \
  -F "p='90'" \
  -F "k_domain='3'"
```

Include a prior `gs1` milestone response:

```{bash dcat-bash-rich, eval=!is_windows}
HOST="https://james.groeidiagrammen.nl"

cat > /tmp/child_rich.json << 'EOF'
{
  "Format": "3.1",
  "clientDetails": [ { "bdsNumber": 20, "value": "20260115" } ],
  "clientMeasurements": [
    { "varName": "gs1moc002", "values": [ { "date": "20260504", "value": "1" } ] }
  ],
  "nestedDetails": []
}
EOF

curl -sX POST "$HOST/dcat/calculate/json?auto_unbox=true" \
  -H "accept: application/json" \
  -F "txt=@/tmp/child_rich.json;type=application/json" \
  -F "population='GSED-NLD'" \
  -F "p='90'" \
  -F "k_domain='3'"
```

### Python

Install the `requests` package once:

```bash
pip install requests
```

POST a local file:

```{python dcat-python-file, eval=FALSE}
import subprocess, requests

host = "https://james.groeidiagrammen.nl"
fn = subprocess.check_output(
    ["Rscript", "-e",
     'cat(system.file("examples", "example_v3.1.json", package = "bdsreader"))'],
    text=True
).strip()

with open(fn, "rb") as f:
    resp = requests.post(
        f"{host}/dcat/calculate/json",
        params={"auto_unbox": "true"},
        files={
            "txt":        ("example_v3.1.json", f, "application/json"),
            "population": (None, "GSED-NLD"),
            "p":          (None, "90"),
            "k_domain":   (None, "3"),
        },
    )
resp.raise_for_status()
print(resp.json())
```


Include a prior `gs1` milestone response:

```{python dcat-python-rich, eval=FALSE}
import requests, json, tempfile, os

host = "https://james.groeidiagrammen.nl"

child_data = {
    "Format": "3.1",
    "clientDetails": [{"bdsNumber": 20, "value": "20260115"}],
    "clientMeasurements": [
        {
            "varName": "gs1moc002",
            "values": [{"date": "20260504", "value": "1"}],
        }
    ],
    "nestedDetails": [],
}

with tempfile.NamedTemporaryFile(suffix=".json", mode="w", delete=False) as f:
    json.dump(child_data, f)
    tmp = f.name

with open(tmp, "rb") as f:
    resp = requests.post(
        f"{host}/dcat/calculate/json",
        params={"auto_unbox": "true"},
        files={
            "txt": (os.path.basename(tmp), f, "application/json"),
            "population": (None, "GSED-NLD"),
            "p": (None, "90"),
            "k_domain": (None, "3"),
        },
    )
os.unlink(tmp)
resp.raise_for_status()
print(resp.json())
```

:::

## Minimal example input

The smallest valid payload contains only the child's date of birth:

```json
{
  "Format": "3.1",
  "clientDetails": [
    { "bdsNumber": 20, "value": "20260115" }
  ]
}
```

A richer payload with prior milestone responses:

```json
{
  "Format": "3.1",
  "clientDetails": [
    { "bdsNumber": 20, "value": "20260115" }
  ],
  "clientMeasurements": [
    {
      "varName": "gs1moc002",
      "values": [ { "date": "20260504", "value": "1" } ]
    },
    {
      "varName": "gs1lgc010",
      "values": [ { "date": "20260504", "value": "0" } ]
    }
  ],
  "nestedDetails": []
}
```

## Resources

- [OpenAPI specification](https://james.groeidiagrammen.nl/docs/)
- [JSON data schema 3.1](https://james.groeidiagrammen.nl/schemas/bds_v3.1.json)
- [Getting started vignette](_getting_started.qmd)
